---
title: "Visualization"
subtitle: "Part I"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

##keep everything else
```



## Scatterplots!!

Create my first scatterplot ever.


```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```


New approach, same plot.

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```



Save and edit a plot object.


```{r}
plot_weather = 
  weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) 

plot_weather + geom_point()
```



## Advanced scatterplot ...
Start with the same one and make it fancy!


```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() + 
  geom_smooth(se = FALSE)
```


What about the 'aes' placement ...?
```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) + 
  geom_smooth()

#the color code only applies to the scatterplot, the color code does not pass on to everywhere
```


Let's facet some things!!

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .2, size = 1) + 
  geom_smooth(se = FALSE, size = 1.5) +
  facet_grid(. ~name)

#what defines rows? (nothing defines the rows here), name difines the columns here
#alpha manipulates the transparency of the dots
```


```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, alpha = tmin, color = name)) +
  geom_point(size = 1) + 
  geom_smooth(se = FALSE, size = 1.5) +
  facet_grid(. ~name)

#alpha transparency is mapped to tmin
```



Let's combine some elements and try a new plot.
```{r}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~name)


##map the dot size to precipitation
```



#Some small notes


How many geoms have to exist?

You can have whatever geoms you want.

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```


You can use a neat geom!

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d() + 
  geom_point(alpha = 0.3)
  
```


## Univariate plots

Histograms are really great.

```{r}
weather_df %>%
  ggplot(aes(x = tmin)) +
  geom_histogram()
```


Can we add color ...
```{r}
weather_df %>%
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
```

```{r}
weather_df %>%
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~name)
```



Let's try a new geometry!


```{r}
weather_df %>%
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3, adjust = .5)
```


What about box plots?

```{r}
weather_df %>%
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot() 
```


Trendy plots :-)

Violin plots!

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

Ridge plots -- the most popular plot of 2017


```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```





## Save and Embed

Let's save a scatterplot.



```{r}
weather_plot = ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("./results/weather_plot.pdf", weather_plot, width = 8, height = 5)

```



What about embedding?

```{r, fig.width=6, fig.height=3}
weather_plot
```


```{r, fig.width=12, fig.height=6}
weather_plot
```