---
title: "Visualization"
subtitle: "Part II"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df

##keep everything else
```


##Remember this plot?





```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(alpha=.5)
```





## labels

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(alpha=.5) +
  labs(
    title = "temperature plot",
    x = "minimum daily temperature (C)",
    y = "minimum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2017."
  )
```




##scales

start with the same plot; x and y scales.
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(alpha=.5) +
  labs(
    title = "temperature plot",
    x = "minimum daily temperature (C)",
    y = "minimum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
scale_x_continuous(
  breaks = c(-15, 0, 15),
  labels = c("-15 C", "0", "15")
) +
scale_y_continuous(
  trans = "sqrt",
  position = "right"
)
```





Look at color scales

```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(alpha=.5) +
  labs(
    title = "temperature plot",
    x = "minimum daily temperature (C)",
    y = "minimum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100,300))
```



```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(alpha=.5) +
  labs(
    title = "temperature plot",
    x = "minimum daily temperature (C)",
    y = "minimum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)
```





##Themes


Shift the legend.
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(alpha=.5) +
  labs(
    title = "temperature plot",
    x = "minimum daily temperature (C)",
    y = "minimum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme(legend.position = "bottom")
```



Change the overall theme.
```{r}
weather_df %>%
  ggplot(aes(x=tmin, y=tmax, color = name)) +
  geom_point(alpha=.5) +
  labs(
    title = "temperature plot",
    x = "minimum daily temperature (C)",
    y = "minimum daily temperature (C)",
    caption = "Data from rnoaa package; temperature in 2017."
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```



## Setting options
```{r, eval=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```




## Data args in 'geom'


```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```



## 'patchwork'


remember faceting?

```{r}
weather_df %>%
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .2, size = 1) + 
  facet_grid(. ~name)
```

what happens when you want multipanel plots but can't facet?

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) + tmax_date_p
(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```




## Data manipulation

control your factors
making change to the order of the category by data manipulation instead of ggplot
```{r}
weather_df %>%
  mutate(
    name = factor(name),
    name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```



What if I wanted densities for tmin and tmax simultaneously?


recognizing that this is also a data tidying issue
```{r}
weather_df %>%
  select(name, tmax, tmin) %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") %>% 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```









